Lab 2 Final Report: COVID-19 Case Rate vs Population Demographics and Mask Policy

### 2020-12-09 Jackson_Liu_Temlock_Zhang_Lab2_Final_Report.Rmd
Final updates to report

 **Initial reassignment of common variables used across models:**
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("GGally")
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
library(dplyr)
library(GGally)
library(car)
```

@@

@@

#### i. I.I.D Assumption

@@

@@

The economic standing of a state not only affects the individual households of its population, but also includes how it interacts with other states and how its peoples’ lives might have been economically changed by the pandemic. One way that a state may increase its number of COVID cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry. In addition, if certain key industries to a state, such as tourism, layed off many workers, they would likely have to take lower paying close contact jobs such as in the service or gig economy in order to continue having a livelihood. This would further lead to a higher COVID case rate.

These effects may be included with a variable that demonstrates how much of a state’s GDP is from tourism. It would be related to other economic indicator variables, such as the “real time” poverty rate during the pandemic. Although poverty is a broader phenomena, the number of people in poverty in the pandemic could be higher due to negative growth of the tourism industry. Therefore, it is expected that states with a greater focus on tourism may also have more people in poverty in the pandemic. While poverty rate is included in the model, it is based off of data from 2018 before any pandemic related effects would have occurred. Assuming real time poverty was included in the model, a GDP percentage of tourism would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive. This would result in a positive omitted variable bias away from zero. Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand a model including only poverty may be an effect of omitted variable bias. Again, because the model in this analysis uses 2018 poverty data, this would not be related to a bias from this omitted variable.

>>>>

 **Initial reassignment of common variables used across models:**
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
library(dplyr)
library(car)
```

@@

Based on the results of the data exploration above, the following equation is used to generate the regression analysis for model 3.

$$
\begin{aligned}
 case\_rate\_100k =\beta_0 &+\beta_1senior\_per\_100k +\beta_2poverty\_per\_100k +\beta_3unemployed\_per\_100k + \beta_4mask\_public\_bool + \beta_5log(population\_density)
\end{aligned}
$$

@@

#### i. IID Assumption

@@

Setting aside the uncertainty provided that the sample is qualified for the IID condition, all three of the models generally meet the CLM assumptions. Therefore, the coefficients generated from the regression analyses are unbiased estimates of the actual relationship investigated in this report. The uncertainty associated with these coefficients is also unbiased. Moreover, it is plausible to apply t-tests based on classical standard errors for these regression coefficients. Thus, the discussions and conclusions drawn from these regression results are assumed to be credible. However, if the dataset was more granular (e.g., case rate at the county-level rather than at the state-level), the IID assumption would be better satisfied. It would lead to greater independence for each data point relative to the rest of the data set. It would also take more advantage of the unbiased estimators so far demonstrated in the model, as a higher number of data points would produce more accurate values for the model coefficients. Overall, improving the source data would lead to greater reliability in the regression work and conclusions.

@@

The economic standing of a state not only affects the individual households of its population, but also includes how it interacts with other states and how its peoples’ lives might have been economically changed by the pandemic. One way that a state may increase its number of COVID cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry. In addition, if certain key industries to a state, such as tourism, laid off many workers, they would likely have to take lower paying close-contact jobs such as in the service industry or gig economy in order to continue to have a livelihood. This would further lead to a higher COVID case rate.

These effects may be included with a variable that demonstrates how much of a state’s GDP is from tourism. It would be related to other economic indicator variables, such as the “real time” poverty rate during the pandemic. Although poverty is a broader phenomena, the number of people in poverty in the pandemic could be higher due to negative growth of the tourism industry. Therefore, it is expected that states with a greater focus on tourism may also have more people in poverty in the pandemic. While poverty rate is included in the model, it is based off of data from 2018 before any pandemic related effects would have occurred. Assuming real time poverty was included in the model, a GDP percentage of tourism would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive. This would result in a positive omitted variable bias away from zero. Without a clear proxy in the dataset, and with the nature of this variable touching on the IID assumption between interaction among states, it should be further investigated whether a model including only poverty may be affected by omitted variable bias. Again, because the model in this analysis uses 2018 poverty data, this would not be related to a bias from this omitted variable.
